plyranges: a grammar for transforming genomics data

### 2018-04-30 paper.Rmd
tidy up of table

@@

*

\begin{table}[ht]
\label{tab:grammar}
\centering
\begin{tabular}{|l|l|p{4cm}|}
  \hline
  & Verb &  Description \\ 
  \hline
  Arithmetic &  {\tt mutate()} & modifies any column \\
  Arithmetic & `select()` & select columns \\
  Arithmetic & `arrange()` & sort by columns \\
  Arithmetic & `set\_width()` & modifies width \\
  Arithmetic & `stretch()` & extend range by fixed amount \\
  Arithmetic & `shift\_(direction)` & shift coordinates \\
  Arithmetic & `flank\_(direction)` & generate flanking regions \\
  Arithmetic & `\%setop\%` & row-wise set operations \\
  Restriction & `filter()` & subset rows \\
  Restriction & `filter\_by\_overlaps()` & subset by overlap \\
  Restriction & `filter\_by\_non\_overlaps()` & subset by no overlap \\
  Merging & `join\_overlap\_()` & merge by overlapping ranges \\
  Merging & `join\_(follow/precedes/nearest)` & merge nearest neighbours \\
  Merging & `(setop)\_ranges` & range-wise set operations \\
  Aggregation & `summarise()` & aggregate over column(s) \\ 
  Aggregation & `disjoin\_ranges()` & aggregate column(s) over the union of end coordinates \\
  Aggregation & `reduce\_ranges()` & aggregate column(s) by merging overlapping and neighbouring ranges \\
  Modifier & `group\_by()` & partition by column(s)  \\ 
  Modifier & `group\_by\_overlaps()` & partition by overlaps \\
  Modifier & `anchor\_direction()` & fix coordinates \\
   \hline
\end{tabular}
\caption{Overview of `plyranges` grammar. Expand here (bold to emphasise
dplyr adaptations)}
\end{table}

>>>>

\begin{table}[ht]
\label{tab:grammar}
\centering
\begin{tabular}{|l|l|p{4cm}|}
  \hline
  & Verb &  Description \\ 
  \hline
   & \textbf{\emph{summarise()}} & aggregate over column(s) \\ 
   Aggregation & \emph{disjoin\_ranges()} & aggregate column(s) over the union of end coordinates \\
   &  \emph{\reduce\_ranges()} & aggregate column(s) by merging overlapping and neighbouring ranges \\
   \hline
   &  \textbf{\emph{mutate()}} & modifies any column \\
   & \textbf{\emph{select()}} & select columns \\
  Arithmetic (Unary) & \textbf{\emph{arrange()}} & sort by columns \\
   & \emph{stretch()} & extend range by fixed amount \\
   &  \emph{shift\_(direction)} & shift coordinates \\
   & \emph{flank\_(direction)} & generate flanking regions \\
   & \emph{\%intersection\% } & row-wise intersection \\
   & \emph{\%union\%} & row-wise union \\
  Arithmetic (Binary) & { \tt \%setdiff\% } & row-wise set difference \\
   & \emph{between()} & row-wise gap range \\
   & \emph{span()} & row-wise spanning range \\
   \hline
    & \emph{join\_overlap\_*()} & merge by overlapping ranges \\
    & \emph{join\_nearest} & merge by nearest neighbour ranges \\
    & \emph{join\_follow} & merge by following ranges \\
    Merging & \emph{join\_precedes} & merge by preceding ranges \\
    & \emph{union_ranges} & range-wise union \\
    & \emph{intersect_ranges} & range-wise intersect \\
    & \emph{setdiff_ranges} & range-wise set difference \\
    & \emph{complement_ranges} & range-wise union \\
  \hline
   & \textbf{\emph{anchor\_direction()}} & fix coordinates at direction \\
  Modifier & \textbf{\emph{group\_by()}} & partition by column(s)  \\ 
   & \emph{group\_by\_overlaps()} & partition by overlaps \\\
   \hline
   & \textbf{\emph{filter()}} & subset rows \\
  Restriction & \emph{filter\_by\_overlaps()} & subset by overlap \\
    & \emph{filter\_by\_non\_overlaps()} & subset by no overlap \\
   \hline
\end{tabular}
\caption{Overview of the \texttt{plyranges} grammar. The core verbs are
briefly described and categorised into one of: aggregation, unary or binary
arithmetic, merging, modifier, or restriction. A verb is given bold text if
its origin is from the \texttt{dplyr} grammar.}
\end{table}

@@
